### 2012-10-25 Docs/Thesis/system_design_and_interface.tex
Full initial draft of system design chapter

>>>>

\subsection{Feedback}
\label{subsection:System_Design_and_Interface-Combinators-Feedback}

It is often useful for a signal function to receive a portion of its
own output as input. This is especially useful when we have two
signal functions which we would like to mutually interact. We cannot
just serially compose them, we must also bring the output of the second
back around to the first. Many signal-processing algorithms also depend
on feedback. The combinator which provides this ability is shown
in Figure~\ref{figure:feedback_combinator}.

\begin{figure}
\begin{code}
loop ::    ((svIn :^: svLoop) :~> (svOut :^: svLoop))
        -> svIn :~> svOut
\end{code}
\hrule
\caption{Feedback combinator.}
\label{figure:feedback_combinator}
\end{figure}

This combinator provides decoupling for signals
(the input signal is the output signal at the previous time-step)
but not events (event occurrences are supplied to the combinator immediately).
This means that the programmer has the responsibility to ensure that feedback
does not generate an infinite sequence of events in a single time-step.

\subsection{Event-Specific Combinators}
\label{subsection:System_Design_and_Interface-Combinators-Event_specific_combinators}

Several combinators are provided for manipulating, suppressing, and generating events.
Output events, in these combinators, are triggered by input events.

Each of the combinators has an option variant and a list variant. The option variant
produces an output event occurrence whenever the application of the supplied function
to the input event produces a value. The list version produces an event occurrence for
each of the elements of the output list. The combinators are shown in Figure~\ref{figure:event_specific_combinators}.

\begin{figure}
\begin{code}
-- Apply the function to each input occurrence,
-- and produce an occurrence for each Just.
filter :: (a -> Maybe b) -> SVEvent a :~> SVEvent b

-- Apply the function to each input occurrence,
-- and produce an occurrence for each list element
filterList :: (a -> [b]) -> SVEvent a :~> SVEvent b

-- Apply the function to the stored accumulator
-- and the event occurrence, replacing the accumulator
-- and possibly outputting an occurrence
accumulate ::    (a -> b -> (Maybe c, a))
              -> a
              -> SVEvent b :~> SVEvent c

-- Apply the function to the stored accumulator
-- and the event occurrence, replacing the
-- accumulator and outputting an event occurrence
-- for each element of the list
accumulateList ::    (a -> b -> ([c]))
                  -> a
                  -> SVEvent b :~> SVEvent c
\end{code}
\hrule
\caption{Event-specific combinators.}
\label{figure:event_specific_combinators}
\end{figure}

The filter combinators are stateless, and thus apply the function to only the new
input value. They are useful for suppressing events, as well as for extracting one
of multiple cases of a datatype. For instance, a splitter for events carrying
{\tt Either}-valued occurrences could be written as:

\begin{code}
getLeft :: Either a b -> Maybe a
getLeft (Left x) = Just x
getLeft _ = Nothing

getRight :: Either a b -> Maybe b
getRight (Right x) = Just x
getRight _ = Nothing

split :: SVEvent (Either a b) :~> (SVEvent a :^: SVEvent b)
split = copy >>> first (filter getLeft) >>> second (filter getRight)
\end{code}

The accumulate combinators are stateful, applying the supplied function
to both the input value and an accumulator. This function has two results:
the option or list of output event occurrence values, and the new value
for the accumulator.

The accumulator is useful when responses to multiple event occurrences
(from one or more sources) must be coordinated. For instance, in the
benchmark application (see Chapter~\ref{chapter:Evaluation_and_Comparisons},
a table is maintained that allows knowledge from previous event occurrences
(packets from a network switch) to be used in deciding where the present
packet ought to go.

\subsection{Joining}
\label{subsection:System_Design_and_Interface-Combinators-Joining}

The joining combinators provide the ability to combine two event
streams, two signals, or a signal and an event stream. These
combinators are shown in Figure~\ref{figure:joining_combinators}

The {\tt union} combinator is a non-deterministic merge of event
streams. Any event which occurs on either input will occur
on the output. For simultaneous event occurrences, the order of occurrence
is not guaranteed, but the occurrence itself is. This construct
is also guaranteed to respect the relationship of event occurrences to sampling
intervals.

The {\tt combineSignals} combinator applies a binary function pointwise to two
signals, and produces the result of this application as a third signal.

The {\tt capture} combinator adds the last-sampled value of a signal at the time
of an event occurrence to that event occurrence.

These three combinators together provide the ability to combine elements of
signal vectors. By combining these combinators, arbitrary signal vectors can
be reduced.

\begin{figure}
\begin{code}
union          :: (SVEvent a :^: SVEvent a) :~> SVEvent a
combineSignals :: (a -> b -> c) -> (SVEvent a :^: SVEvent b) :~> SVEvent c
capture        :: (SVSignal a :^: SVEvent b) :~> SVEvent (b, a)
\end{code}
\hrule
\caption{Joining combinators.}
\label{figure:joining_combinators}
\end{figure}
